---
title: "Assignment 2"
author: 'Bodireddy Vikas Reddy netID: hg8118'
date: "2023-03-01"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  **(Exercise 4.4) Consider an experiment taste-testing six types of
    chocolate chip cookies: 1 (brand A, chewy, expensive), 2 (brand A,
    crispy, expensive), 3 (brand B, chewy, inexpensive), 4 (brand B,
    crispy, inexpensive), 5 (brand C, chewy, expensive), and 6 (brand D,
    crispy, inexpensive). We will use twenty different raters randomly
    assigned to each type (120 total raters).**

    (brand A, chewy, expensive)

    (brand A, crispy, expensive)

    (brand B, chewy, inexpensive)

    (brand B, crispy, inexpensive)

    (brand C, chewy, expensive)

    (brand D, crispy, inexpensive)

    **(a) Design contrasts to compare chewy with crispy, and expensive
    with inexpensive.**

    Answers:

    The Null hypothesis for the above **chewy with crispy** is

    $$H_0 : (\mu_1+\mu_2+\mu_5)= (\mu_2+\mu_4+\mu_6)$$

    ans alternate hypotheis is

    $$H_A: (\mu_1+\mu_2+\mu_5) \not= (\mu_2+\mu_4+\mu_6)$$

    and the conntrasts for the chewy and crispy is : $w$ **=
    (1,-1,1,-1,1,-1)**

    The null hypothesis for **expensive and inexpensive** is

    $$H_0 : (\\mu_1+\\mu_2+\\mu_5)= (\\mu_3+\\mu_4+\\mu_6)$$

    $$H_A : (\\mu_1+\\mu_2+\\mu_5)\not= (\\mu_3+\\mu_4+\\mu_6)$$

    The contrasts to compare expensive with inexpensive : $w^*$**=
    (1,1,-1,-1,1,-1)**

    **(b) Are your contrasts in part (a) orthogonal? Why or why not?**

    Answers: no the contrasts are not orthogonal, as contrasts are said
    to be orthogonal if

    $$\sum_{i=1}^{g}w_i w_i^*/n_i = 0 $$

    Here, $$\sum_{i=1}^{g}w_i w_i^*/n_i =
    (1*1)+(-1*1)+(1*-1)+(-1*-1)+(1*1)+(-1*-1) /20$$

    $$\sum_{i=1}^{g}w_i w_i^*/n_i = 1-1-1+1+1+1 /20$$

    $$\sum_{i=1}^{g}w_i w_i^*/n_i = 2/20 != 0$$

**2. (Problem 4.1) A consumer testing agency obtains four cars from each
of six makes: Ford, Chevrolet, Nissan, Lincoln, Cadillac, and Mercedes.
Makes 3 and 6 are imported while the others are domestic; makes 4, 5,
and 6 are expensive while 1, 2, and 3 are less expensive; 1 and 4 are
Ford products, while 2 and 5 are GM products. We wish to compare the six
makes on their oil use per 100,000 miles driven. Dataset is uploaded to
Canvas with name car.dat. Use the following cmd to read data into R,
read.table("car.dat", header = TRUE) where the car markers are coded
from 1 to 6 corresponding to Ford, Chevrolet, Nissan, Lincoln, Cadillac,
and Mercedes, respectively. Data are recorded in the response
variable.** **(a) Compute the Analysis of Variance table for this
experiment. What would you conclude?**

```{r}
car_data <- read.table("C:\\Users\\STSC\\Desktop\\631\\assignments\\Assignment 2\\car.dat", 
                       header = TRUE)
head(car_data)

```

```{r}
#full model
car_data$ftype <- as.factor(car_data$type)
car_fit <-aov(response ~ ftype , data = car_data)
anova(car_fit)

```

**Answer: by performing anova to the above dataset we can see that there
is at least one mean difference between the given treatments. as** **P
Value is (4.887e-08 \<** $\alpha$ **= 0.05) hence we reject the null
hypothesis** $H_0$

**(b) There are many meaningful contrasts that can be constructed.
Consider two of them below. Design the contrasts to answer the following
research questions and obtain the corresponding 95% confidence interval
for two-sided alternatives. Report the meaning of confidence interval
for each contrast.**

```{r}
library(emmeans)

lsmtype_1 <- lsmeans(car_fit, ~ftype)
lsmtype_1
```

**i.Design a contrast to compare the oil use in imported cars with that
in exported cars.**

**Answers: Since the confidence interval range does not have 0 in
between the upper and lower limits(-3.73,-2.02), we reject the null
hypothesis** $H_0$

```{r}
contrast_1 = list("imported cars with respect to exported cars" =c(1,1,-2,1,1,-2))
summary(contrast(lsmtype_1,contrast_1),infer = c(T,T),level= 0.95, side = "Two-sided ")

```

**ii. Design a contrast to compare the oil use in cheap cars with that
in expensive cars.**

**Answers: Since the confidence interval range does not have 0 in
between the upper and lower limits(-2.9,-1.7), we reject the null
hypotheis** $H_0$

```{r}
lsmtype_2 <- lsmeans(car_fit, ~ ftype)
contrast_2 <- list("cheap cars with repect to expensive cars " = c(1,1,1,-1,-1,-1))
summary(contrast(lsmtype_2,contrast_2), infer = c(T,T), level = 0.95, side = "Two-sided")
```
